---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)

# Answer:
print("There are total of 32 variables and 11 observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are 7 discrete variables and 4 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- colMeans(mtcars["mpg"])
v <- var(mtcars["mpg"])
s <- sd(as.numeric(unlist(mtcars["mpg"])))

# 
print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ",  v, " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
library(tidyverse)
library(magrittr)
mtcars %>%
  group_by(cyl) %>%
  summarize(AvgMpg = mean (mpg))

mtcars %>%
  group_by(gear) %>%
  summarize(SdMpg = sd(as.numeric(unlist(mpg))))


```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
mtcars %>% 
 select(cyl, gear) %>% 
 table()

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
g<-PlantGrowth

Gp1<-g$weight[1:10]
Gp2<-g$weight[11:20]
Gp3<-g$weight[21:30]
boxplot(Gp1,Gp2,Gp3,names=c("Cntrl","Trt1","Trt2"), main="weight of the plant seperated by group", xlab="groups of the plant", ylab="weight of the plant")

```

Result:

=> Report a paragraph to summarize your findings from the plot!
The group of Trt2 is the heaviest weighing 5.5. Secondly, the lowest weighing in the group is Trt1 at the weight of 4.5. Lastly, Cntrl is the median of the group, weighing in about 5.1.


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg, breaks=10, xlab="Mile per gallon", main="Histogram of cars classes by miles per gallon")


print("Most of the cars in this data set are in the class of 15 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
plot(USArrests)
plot(USArrests$Murder,USArrests$Assault,xlab='Murder',ylab='Assault', main="Murder vs Assault")

```

Result:

=> Report a paragraph to summarize your findings from the plot!
The safest out of the 6 states would be Colorado. On the other hand, the most dangerous state out of the six would be Alaska. There are a lot more cases of assault on these states.


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
m1 <- colMeans(housingData["Market.Value.per.SqFt"])
v1 <- var(housingData["Market.Value.per.SqFt"])
s1 <- sd(as.numeric(unlist(housingData["Market.Value.per.SqFt"])))

print(paste("The average of Market.Value.per.SqFt from this data set is ", m1 , " with variance ",  v1, " and standard deviation", s1 , "."))

m2<- colMeans(housingData["Year.Built"])
v2<- var(housingData["Year.Built"])
s2<- sd(as.numeric(unlist(housingData["Year.Built"])))

print(paste("The average of Year.Built from this data set is ", m2, " with variance ",  v2," and standard deviation", s2, "."))

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
plot(housingData$Year.Built,housingData$Market.Value.per.SqFt,xlab='Year.Built',ylab='Market.Value.per.SqFt', main="Year.Built vs Market.Value.per.SqFt")
hist(housingData$Year.Built, xlab="MYear.Built", main="Histogram of number of houses by Year.Built")
```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!
The market value per sqft. from the years 1900 to 1950 are much less than the year 1950 to 2000 by a significant amount. It would obviously be like that as the market would prefer something new than older.